Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-04-08 WSSSPE4_report/WSSSPE4_report.tex
another affil update
adding last missing affiliation
additions to abstract and conclusions

\author{Daniel~S.~Katz$^{(1)}$,
Kyle~E.~Niemeyer$^{(2)}$,
Sandra~Gesing$^{(3)}$,
Lorraine~Hwang$^{(4)}$,
Simon~Hettrick$^{(5)}$,
Alice~Allen$^{(6)}$,
Jean~Salac$^{(7)}$,
Neil~Chue~Hong$^{(8)}$,
Santiago~Nunez-Corrales$^{(9)}$
Anshu~Dubey$^{(10)}$,
R.~Stuart~Geiger$^{(11)}$,
Jonah~Miller$^{(12)}$,
Emily~Chen$^{(13)}$,
Wolfgang~Bangerth$^{(14)}$,
\\Author order to be determined based on contributions
}

%
\thanks{{}$^{(1)}$ National Center for Supercomputing Applications (NCSA) \&
Department of Computer Science  \&
Department of Electrical and Computer Engineering  \&
School of Information Sciences (iSchool),
University of Illinois, Urbana--Champaign, IL, USA; d.katz@ieee.org; ORCID: 0000-0001-5934-7525}
%
\thanks{{}$^{(2)}$ School of Mechanical, Industrial, and Manufacturing Engineering,
Oregon State University, Corvallis, OR, USA; kyle.niemeyer@oregonstate.edu; ORCID: 0000-0003-4425-7097}
%
\thanks{{}$^{(3)}$ Center for Research Computing \& Department of Computer Science and Engineering,
University of Notre Dame, IN, USA; sandra.gesing@nd.edu; ORCID: 0000-0002-6051-0673}
%
\thanks{{}$^{(4)}$ University of California, Davis, CA, USA; ljhwang@ucdavis.edu; ORCID: 0000-0002-1021-3101}
%
\thanks{{}$^{(5)}$ The Software Sustainability Institute, Southampton, UK; s.hettrick@software.ac.uk; ORCID: 0000-0002-6809-5195}
%
\thanks{{}$^{(6)}$ Astrophysics Source Code Library, Astronomy Department, University of Maryland, College Park, MD, USA; aallen@ascl.net; ORCID: 0000-0003-3477-2845}
%
\thanks{{}$^{(7)}$ \note{info needed}}
%
\thanks{{}$^{(8)}$ \note{info needed}}
%
\thanks{{}$^{(9)}$ Illinois Informatics Institute (I3) and National Center for Supercomputing Applications (NCSA), University of Illinois at Urbana-Champaign, IL, USA;
nunezco2@illinois.edu;
ORCID: 0000-0003-4342-6223}
%
\thanks{{}$^{(10)}$ Mathematics and Computer Science Division,
Argonne National Laboratory,
Lemont, IL, USA \& 
University of Chicago, 
Chicago, IL, USA;
Adubey@anl.gov;
ORCID: 0000-0003-3299-7426}
%
\thanks{{}$^{(11)}$ University of California, Berkeley, Berkeley Institute for Data Science (BIDS), Berkeley, CA, USA; stuart@stuartgeiger.com; ORCID: 0000-0001-7215-0532}
%
\thanks{{}$^{(12)}$ Perimeter Institute for Theoretical Physics \&
Department of Physics, University of Guelph, 
Guelph, ON, Canada; 
jmiller@perimeterinstitute.ca; ORCID: 0000-0001-6432-7860}
%
\thanks{{}$^{(13)}$ University of Illinois at Urbana-Champaign, USA; echen35@illinois.edu; ORCID: 0000-0003-2578-0296}
%
\thanks{{}$^{(14)}$ Department of Mathematics, Colorado State
  University, Fort Collins, CO, USA; bangerth@colostate.edu; ORCID: 0000-0003-2311-9402}

\begin{abstract}
This report records and discusses the Fourth Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE4). The report
includes a description of the keynote presentation of the workshop,
which % TODO
% It also summarizes a set of lightning
% talks in which speakers highlighted to-the-point lessons and challenges
% pertaining to sustaining scientific software.
% The final and main contribution of the report is a summary of the
% discussions, future steps, and future organization for a set of self-organized
% working groups on topics including developing pathways to funding scientific
% software; constructing useful common metrics for crediting software
% stakeholders; identifying principles for sustainable software engineering
% design; reaching out to research software organizations around the world; and
% building communities for software sustainability. For each group, we include a
% point of contact and a landing page that can be used by those who want to join
% that group's future activities. The main challenge left by the workshop is to
% see if the groups will execute these activities that they have scheduled, and
% how the WSSSPE community can encourage this to happen.

@@

This report is based on the events at the workshop and the submitted materials.
There events included discussion of the WSSSPE mission and vision, a keynote,
a set of presentations, a panel, and a number of working groups.  About a day of the
workshop was spent with participants in the working groups, which occurred in parallel
with each other.

@@

In WSSSPE4, we

\todo{discuss Twitter activity}

>>>>

\author{Daniel~S.~Katz$^{(1)}$,
Kyle~E.~Niemeyer$^{(2)}$,
Sandra~Gesing$^{(3)}$,
Lorraine~Hwang$^{(4)}$,
Simon~Hettrick$^{(5)}$,
Alice~Allen$^{(6)}$,
Jean~Salac$^{(7)}$,
Neil~Chue~Hong$^{(8)}$,
Santiago~Nunez-Corrales$^{(9)}$
Anshu~Dubey$^{(10)}$,
R.~Stuart~Geiger$^{(11)}$,
Jonah~Miller$^{(12)}$,
Emily~Chen$^{(13)}$,
Wolfgang~Bangerth$^{(14)}$%,
%\\Author order to be determined based on contributions
}

%
\thanks{{}$^{(1)}$ National Center for Supercomputing Applications (NCSA) \&
Department of Computer Science  \&
Department of Electrical and Computer Engineering  \&
School of Information Sciences (iSchool),
University of Illinois, Urbana--Champaign, IL, USA; d.katz@ieee.org; ORCID: 0000-0001-5934-7525}
%
\thanks{{}$^{(2)}$ School of Mechanical, Industrial, and Manufacturing Engineering,
Oregon State University, Corvallis, OR, USA; kyle.niemeyer@oregonstate.edu; ORCID: 0000-0003-4425-7097}
%
\thanks{{}$^{(3)}$ Center for Research Computing \& Department of Computer Science and Engineering,
University of Notre Dame, IN, USA; sandra.gesing@nd.edu; ORCID: 0000-0002-6051-0673}
%
\thanks{{}$^{(4)}$ University of California, Davis, CA, USA; ljhwang@ucdavis.edu; ORCID: 0000-0002-1021-3101}
%
\thanks{{}$^{(5)}$ The Software Sustainability Institute, Southampton, UK; s.hettrick@software.ac.uk; ORCID: 0000-0002-6809-5195}
%
\thanks{{}$^{(6)}$ Astrophysics Source Code Library, Astronomy Department, University of Maryland, College Park, MD, USA; aallen@ascl.net; ORCID: 0000-0003-3477-2845}
%
\thanks{{}$^{(7)}$ University of Virginia, Charlottesville, VA, USA; jeansalac@virginia.edu; ORCID: 0000-0001-5971-9333}
%
\thanks{{}$^{(8)}$ University of Edinburgh, UK; N.ChueHong@epcc.ed.ac.uk;
ORCID: 0000-0002-8876-7606}
%
\thanks{{}$^{(9)}$ Illinois Informatics Institute (I3) and National Center for Supercomputing Applications (NCSA), University of Illinois at Urbana-Champaign, IL, USA;
nunezco2@illinois.edu;
ORCID: 0000-0003-4342-6223}
%
\thanks{{}$^{(10)}$ Mathematics and Computer Science Division,
Argonne National Laboratory,
Lemont, IL, USA \& 
University of Chicago, 
Chicago, IL, USA;
Adubey@anl.gov;
ORCID: 0000-0003-3299-7426}
%
\thanks{{}$^{(11)}$ University of California, Berkeley, Berkeley Institute for Data Science (BIDS), Berkeley, CA, USA; stuart@stuartgeiger.com; ORCID: 0000-0001-7215-0532}
%
\thanks{{}$^{(12)}$ Perimeter Institute for Theoretical Physics \&
Department of Physics, University of Guelph, 
Guelph, ON, Canada; 
jmiller@perimeterinstitute.ca; ORCID: 0000-0001-6432-7860}
%
\thanks{{}$^{(13)}$ University of Illinois at Urbana-Champaign, USA; echen35@illinois.edu; ORCID: 0000-0003-2578-0296}
%
\thanks{{}$^{(14)}$ Department of Mathematics, Colorado State
  University, Fort Collins, CO, USA; bangerth@colostate.edu; ORCID: 0000-0003-2311-9402}

\begin{abstract}
This report records and discusses the Fourth Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE4). The report
includes a description of the keynote presentation of the workshop, the
mission and vision statements that were drafted at the workshop and 
finalized shortly after it, a set of idea papers, position papers, experience
papers, demos, and lightning talks, and a panel discussion.  The main part of the report
covers the set of working groups that formed during the meeting, and for each,
discusses the participant, the objective and goal, and how the objective can
be reached, along with contact information for readers who may want to join
the group. Finally, we present results from a survey of the workshop attendees.

% It also summarizes a set of lightning
% talks in which speakers highlighted to-the-point lessons and challenges
% pertaining to sustaining scientific software.
% The final and main contribution of the report is a summary of the
% discussions, future steps, and future organization for a set of self-organized
% working groups on topics including developing pathways to funding scientific
% software; constructing useful common metrics for crediting software
% stakeholders; identifying principles for sustainable software engineering
% design; reaching out to research software organizations around the world; and
% building communities for software sustainability. For each group, we include a
% point of contact and a landing page that can be used by those who want to join
% that group's future activities. The main challenge left by the workshop is to
% see if the groups will execute these activities that they have scheduled, and
% how the WSSSPE community can encourage this to happen.

@@

This report is based on the events at the workshop and the submitted materials.
There events included discussion of the WSSSPE mission and vision, a keynote,
a set of presentations, a panel, and a number of working groups.  One full day of the
workshop was spent with participants in the working groups, which occurred in parallel
with each other.  And each of the working groups left with a plan for how they could move
forward.

@@

In WSSSPE4, we heard about a number of interesting projects and ideas, and used those
ideas to create and form working groups, intended to start at the workshop and then
continue afterwards, to address challenges that arose from the workshop ideas.  This workshop
has reinforced the lesson from WSSSPE3 that it is relatively easy to get motivated people
to attend a meeting and productively spend their time there both doing work and planning
more work, but it is very hard to get that additional work after the meeting to take place.
The main problem seems to be one of time.  Once the attendees have agree to spend
their time at the workshop, they put their energy into doing so productively, but they haven't
really committed themselves to anything more than this, so their energy and effort trails
off, as all of their other commitments (particularly, those they are funded to do as part of
their jobs) come back to the fore. Without a process to resolve this concern, the utility of
further multi-day WSSSPE workshops is unclear.

%\todo{discuss Twitter activity}
